--- 
title: "May the 4th "
subtitle: "SUBTITLE_OF_YOUR_PROJECT<br><br>" 
author: 
  - "AUTHOR_1"
  - "AUTHOR_2"
  - <br>
date: "`r Sys.Date()`"
format: 
  html
--- 

```{r setup, include = F}
library(tidyverse)
library(skimr)   # a better summary of data.frame
library(scales)  # scales for ggplot
library(ggthemes)  # additional ggplot themes
library(socviz)
theme_set(theme_minimal()) # setting the minimal theme for ggplot
# setting default chunk options
knitr::opts_chunk$set(echo = T, eval = T,
                      message=F, warning = F) 
```


```{r}
# library(tidyverse)
# oj <- read_csv('https://bcdanl.github.io/dominick_oj.csv')
# nvars <- format(round(ncol(oj), 0), 
#                 nsmall=0, 
#                 big.mark=",") 
# nobs <- format(round(nrow(oj), 0), 
 #                nsmall=0, 
  #               big.mark=",")


# The number of variables is `r nvars`; the number of observations is `r nobs`
``` 





<br>

**Roses** are <span style="color: red;">red</span>, *violets* are <span style="color: blue;">blue</span>.


# Introduction 
- Why does your research project matter? 

- Motivate your project.

- Summarize the project in the last paragraph.

# Data

Summary statistics(Use `skimr::skim()`)

# Conceptual Framework 

- Justify your ML model.
 - Why the model considers the variables in the model. 

# Model 

- Add some model equation, for example, $y_{i} = \beta_{1} * x_{1i} + \beta_{2} * x_{1i}^{2} + \epsilon_{i}$


- For model equation, better to use `$$` for LaTeX math equations 
$$y_{i} = \beta_{1} * x_{1i} + \beta_{2} * x_{1i}^{2} + \epsilon_{i}$$
- In latex equation, you can add horizontal space by using the following: 
- `\,`
- `\;`
- `quad`
- `qquad`


# Result 

- Summarize the main result of the model you want to focus.

# Discussion

- Discuss the implication of your model result. 

- You can suggest policy suggestion or business strategy. 

# Conclusion 